---
title: "Untitled"
author: "K. J. McKenzie"
date: "28/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("C:/Users/kelse/Documents/Masters R")
```

```{r}
library(tidyverse)
library(readxl)
```

```{r}
df <- read_xlsx("IMG-JGI.xlsx")
df$`Temperature Range` <- as.factor(as.character(df$`Temperature Range`))
df$Phylum <- as.factor(as.character(df$Phylum))
df <- na.omit(df)
table(df$`Temperature Range`)
table(df$Phylum)
df<-df[!(df$`Temperature Range`=="Undefined" | df$`Temperature Range`=="Psychrotrophic"),]
df<-df[!(df$Phylum=="Armatimonadetes" | df$Phylum=="Caldiserica" | df$Phylum=="Candidatus Saccharibacteria" | df$Phylum=="Dictyoglomi" | df$Phylum=="Ignavibacteriae" | df$Phylum=="Calditrichaeota" | df$Phylum=="Chrysiogenetes" | df$Phylum=="Elusimicrobia" | df$Phylum=="Lentisphaerae" | df$Phylum=="unclassified" | df$Phylum=="Candidatus Poribacteria" | df$Phylum=="Coprothermobacterota" | df$Phylum=="Fibrobacteres" | df$Phylum=="Gemmatimonadetes"),]

df$`Temperature Range` <- as.factor(as.character(df$`Temperature Range`))
df$Phylum <- as.factor(as.character(df$Phylum))
table(df$Phylum, df$`Temperature Range`)

table(df$Phylum)

```
Ok so we have:
HT: 26
M: 8838
P: 80
Pt: 126
Ptr: 23
T: 335
Tt: 28
U: 5

Removed all psychroptrophic because that is ambiguous and all undefined


Now I want to extract all phyla with less than 60 genomes to keep, all genomes from the big group with any form of thermo/psychro
```{r}
semidone <- subset(df, Phylum == "Acidobacteria" | Phylum == "Balneolaeota" | Phylum=="Fusobacteria" | Phylum=="Thermotogae" | Phylum =="Chlamydiae" | Phylum =="Deferribacteres" | Phylum=="Nitrospirae" | Phylum=="Synergistetes" | Phylum =="Verrucomicrobia" | Phylum =="Aquificae" | Phylum=="Chlorobi" | Phylum=="Deinococcus-Thermus" | Phylum=="Planctomycetes" | Phylum=="Chloroflexi" | Phylum =="Thermodesulfobacteria")

df <- df[!(df$Phylum == "Acidobacteria" | df$Phylum == "Balneolaeota" | df$Phylum=="Fusobacteria" | df$Phylum=="Thermotogae" | df$Phylum =="Chlamydiae" | df$Phylum =="Deferribacteres" | df$Phylum=="Nitrospirae" | df$Phylum=="Synergistetes" | df$Phylum =="Verrucomicrobia" | df$Phylum =="Aquificae" | df$Phylum=="Chlorobi" | df$Phylum=="Deinococcus-Thermus" | df$Phylum=="Planctomycetes" | df$Phylum=="Chloroflexi" | df$Phylum =="Thermodesulfobacteria"),]
df$`Temperature Range` <- as.factor(as.character(df$`Temperature Range`))
df$Phylum <- as.factor(as.character(df$Phylum))
table(df$Phylum, df$`Temperature Range`)

semidonet <- subset(df, `Temperature Range` =="Hyperthermophile" | `Temperature Range` =="Psychrophile" | `Temperature Range` =="Psychrotolerant" | `Temperature Range` =="Thermophile" | `Temperature Range` =="Thermotolerant")
df <- df[!(df$`Temperature Range`=="Hyperthermophile" | df$`Temperature Range` =="Psychrophile" | df$`Temperature Range` =="Psychrotolerant" | df$`Temperature Range` =="Thermophile" | df$`Temperature Range` =="Thermotolerant"),]
df$`Temperature Range` <- as.factor(as.character(df$`Temperature Range`))
df$Phylum <- as.factor(as.character(df$Phylum))
table(df$Phylum, df$`Temperature Range`)
```

NOw to do the nitty gritty subsetting from the over abundant phyla
First, look at cyano, sprio and tene
Cyano has Glorobacteria 1 and unclassified 139. Hence, random subset of 75
Spirochaetes has 1 class, random subset of 75
Tenericutes has 1 class, random subset of 75

```{r}
cya <- subset(df, Phylum=="Cyanobacteria")
table(cya$Class)
cya <- cya[sample(1:nrow(cya),75, replace=FALSE),]


spi <- subset(df, Phylum=="Spirochaetes")
table(spi$Class)
spi <- spi[sample(1:nrow(spi),75, replace=FALSE),]

ten <- subset(df, Phylum=="Tenericutes")
table(ten$Class)
ten <- ten[sample(1:nrow(ten),75, replace=FALSE),]
```
And now the big bois
Actinobacteria, take all from low 4 classes and 25 from Actinomycetia
Bacteroidetes, take 25 from all and all low
Firmicutes, take 25 from all and all low
Proteobacteria, take 25 from all and all low
```{r}
act <- subset(df, Phylum=="Actinobacteria")
table(act$Class)
act1 <- subset(act, Class=="Acidimicrobiia" | Class=="Nitriliruptoria" | Class=="Thermoleophilia")
act2 <- subset(act, Class=="Coriobacteriia")
act2 <- act2[sample(1:nrow(act2),25, replace=FALSE),]
act3 <- subset(act, Class=="Actinomycetia")
act3 <- act3[sample(1:nrow(act3),25, replace=FALSE),]  
actf <- rbind(act1, act2)
actf <- rbind(actf, act3)
  
Bact <- subset(df, Phylum=="Bacteroidetes")
table(Bact$Class)
bact1 <- subset(Bact, Class=="Chitinophagia" | Class=="Saprospiria" | Class=="unclassified")
bact2 <- subset(Bact, Class=="Bacteroidia")
bact2 <- bact2[sample(1:nrow(bact2),25, replace=FALSE),]
bact3 <- subset(Bact, Class=="Cytophagia")
bact3 <- bact3[sample(1:nrow(bact3),25, replace=FALSE),]
bact4 <- subset(Bact, Class=="Flavobacteriia")
bact4 <- bact4[sample(1:nrow(bact4),25, replace=FALSE),]
bact5 <- subset(Bact, Class=="Sphingobacteriia")
bact5 <- bact5[sample(1:nrow(bact5),25, replace=FALSE),]
bactf <- rbind(bact1, bact2)
bactf <- rbind(bactf, bact3)
bactf <- rbind(bactf, bact4)
bactf <- rbind(bactf, bact5)


Firm <- subset(df, Phylum=="Firmicutes")
table(Firm$Class)
firm1 <- subset(Firm, Class=="unclassified")
firm2 <- subset(Firm, Class=="Bacilli")
firm2 <- firm2[sample(1:nrow(firm2),25, replace=FALSE),]
firm3 <- subset(Firm, Class=="Clostridia")
firm3 <- firm3[sample(1:nrow(firm3),25, replace=FALSE),]
firm4 <- subset(Firm, Class=="Erysipelotrichia")
firm4 <- firm4[sample(1:nrow(firm4),25, replace=FALSE),]
firm5 <- subset(Firm, Class=="Negativicutes")
firm5 <- firm5[sample(1:nrow(firm5),25, replace=FALSE),]
firm6 <- subset(Firm, Class=="Tissierellia")
firm6 <- firm6[sample(1:nrow(firm6),25, replace=FALSE),]
firmf <- rbind(firm1, firm2)
firmf <- rbind(firmf, firm3)
firmf <- rbind(firmf, firm4)
firmf <- rbind(firmf, firm5)
firmf <- rbind(firmf, firm6)

Prot <- subset(df, Phylum=="Proteobacteria")
table(Prot$Class)
prot1 <- subset(Prot, Class=="Acidithiobacillia" | Class=="Oligoflexia" | Class=="Zetaproteobacteria")
prot2 <- subset(Prot, Class=="Alphaproteobacteria")
prot2 <- prot2[sample(1:nrow(prot2),25, replace=FALSE),]
prot3 <- subset(Prot, Class=="Betaproteobacteria")
prot3 <- prot3[sample(1:nrow(prot3),25, replace=FALSE),]
prot4 <- subset(Prot, Class=="Deltaproteobacteria")
prot4 <- prot4[sample(1:nrow(prot4),25, replace=FALSE),]
prot5 <- subset(Prot, Class=="Epsilonproteobacteria")
prot5 <- prot5[sample(1:nrow(prot5),25, replace=FALSE),]
prot6 <- subset(Prot, Class=="Gammaproteobacteria")
prot6 <- prot6[sample(1:nrow(prot6),25, replace=FALSE),]
protf <- rbind(prot1, prot2)
protf <- rbind(protf, prot3)
protf <- rbind(protf, prot4)
protf <- rbind(protf, prot5)
protf <- rbind(protf, prot6)

collection <- rbind(semidone, semidonet)
collection <- rbind(collection, cya)
collection <- rbind(collection, spi)
collection <- rbind(collection, ten)
collection <- rbind(collection, actf)
collection <- rbind(collection, bactf)
collection <- rbind(collection, firmf)
collection <- rbind(collection, protf)
table(collection$`Temperature Range`)
table(collection$`Temperature Range`, collection$Phylum)

write_xlsx(collection, "JGI_final.xlsx")
```
There are only 330 archaea that are going to work properly for this project so I will just keep them all and not do any filtering (446 after filter for no temp detail, 330 after filter for GCA_)
```{r}
archaea <- read_xlsx("archaea.xlsx")
archaea <- archaea %>% drop_na(`Temperature Range`)
archaea <- archaea %>% drop_na(Phylum)
archaea <- archaea %>% drop_na(`NCBI Assembly Accession`)
write_xlsx(archaea, "archaeafiltered.xlsx")
```


Then save whatever column you want at .txt file and then use https://www.ncbi.nlm.nih.gov/sites/batchentrez
to extract what information you want
